Scheduling and Vectorization for MPC

### 2022-03-02 paper_SIMD.tex
Edits to Arrays section. Still need to double check some.

In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle differently def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable. The definition can be an update \texttt{A\_2 = update(A\_1,i,...)} or a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc.. Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars. In addition, array variables are different from scalar variables. Raising and dropping dimension still apply, however, the dimension cannot be raised or dropped to a dimension lower than the canonical one:

@@

The canonical dimensionality of \texttt{res} is 1. Also, the phi-node \texttt{res2 = PHI(res1, res3)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$). After Phase 1, Basic vectorization produces the following (everything is implicitly vectorized along $i$ and $j$ and the update is turned into a regular write):

{\small
\begin{verbatim}
A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
I = raise_dim(i, ((i:num_bins),(j:N)))
for i in range(0, num_bins):
   res1 = PHI(res, res2) // target-less, no drop_dim
   res1' = raise_dim(res1, (j:N))
   for j in range(0, N):
      res2 = PHI(res1', res3)
      tmp1 = (A1 == I)
      tmp2 = (res2 + B1)
      tmp3 = MUX(tmp1, res2, tmp2)
      res3 = res2 // copy propagation: res2 into res3
      res3 = tmp3 // "normal": res3[I,J] = tmp3[I,J]    
return res1
\end{verbatim}
}

Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop. Processing the outer loop is interesting. This is because the PHI node is a target-less node, and therefore, there are no closures! Thus everything can be vectorized along the $i$ dimension:

@@

res1' = raise_dim(res, (j:N)) // raising res to a 2-dim. array
for j in range(0, N):
   res2 = PHI(res1', res3)
   tmp2[I,j] = (res2[I,j] + B1[I,j])
   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
   res3 = res2
   res3[I,j] = tmp3[I,j]
res1 = drop_dim(res2) // Have to add drop_dim before removing redundant phi-node      
return res1
\end{verbatim}
}

The cleanup phase can then get rid of $j$ in the res variables and tmp variables, so the loop operates over the num\_bins-size vectors. 
This will lead to dropping raise\_dim and drop\_dim as well:

@@

for j in range(0, N):
   res2 = PHI(res1', res3) // now 1-dim array
   tmp2[I] = (res2[I] + B1[I,j])
   tmp3[I] = MUX(tmp1[I,j], res2[I], tmp2[I])
   res3 = res2
   res3[I] = tmp3[I]
return res2
\end{verbatim}
}

>>>>

In addition to the changes for the handling of target-less phi-nodes, Basic Vectorization has to handle differently def-use edges $X \rightarrow Y$ where $X$ defines and $Y$ uses an array variable. The definition can be an update \texttt{A\_2 = update(A\_1,i,...)} or a pseudo $\phi$-node \texttt{A\_2 = PHI(A\_0,A\_1)}, etc.. Note that $\phi$-nodes for arrays have no subscript operations the way there are subscript operations in analysis-introduced arrays representing scalars. Raising and dropping dimension still apply, however, the dimension cannot be raised or dropped to a dimension lower than the canonical one:

@@

The canonical dimensionality of \texttt{res} is 1. Also, the phi-node \texttt{res1 = PHI(res, res2)} is a target-less phi-node (the implication being that the inner for loop can be vectorized across $i$). After Phase 1, Basic vectorization produces the following code (statements are implicitly vectorized along $i$ and $j$). \ana{Define the interpretation of vectorized Update!}

{\small
\begin{verbatim}
A1 = raise_dim(A, j, ((i:num_bins),(j:N)))
B1 = raise_dim(B, j, ((i:num_bins),(j:N)))
I = raise_dim(i, ((i:num_bins),(j:N)))
for i in range(0, num_bins):
   res1 = PHI(res, res2) // target-less, no drop_dim
   res1' = raise_dim(res1, (j:N))
   for j in range(0, N):
      res2 = PHI(res1', res3)
      tmp1 = (A1 == I)
      tmp2 = (res2 + B1)
      tmp3 = MUX(tmp1, res2, tmp2)
      res3 = Update(res2, (I,J), tmp3) 
return res1
\end{verbatim}
}

Processing the inner loop in Phase 2 vectorizes \texttt{tmp1 = (A1 == I)} along the $j$ dimension but leaves the rest of the statements in a MOTION loop. Processing the outer loop is interesting. This is because the PHI node is a target-less node, and therefore, there are no closures! Thus, everything can be vectorized along the $i$ dimension:

@@

res1' = raise_dim(res, (j:N)) // raising res to a 2-dim. array
for j in range(0, N):
   res2 = PHI(res1', res3)
   tmp2[I,j] = (res2[I,j] + B1[I,j])
   tmp3[I,j] = MUX(tmp1[I,j], res2[I,j], tmp2[I,j])
   res3 = Update(res2, (I,j), tmp3)
res1 = drop_dim(res2) // Add drop_dim before removing redundant phi-node!      
return res1
\end{verbatim}
}

The cleanup phase can then get rid of $j$ in the \texttt{res} variables and \texttt{tmp} variables, so the loop operates over the \texttt{num\_bins}-size vectors. 
This will lead to dropping raise\_dim and drop\_dim as well:

@@

for j in range(0, N):
   res2 = PHI(res1, res3) // now 1-dim array
   tmp2[I] = ADD_SIMD(res2[I],B1[I,j])
   tmp3[I] = MUX_SIMD(tmp1[I,j], res2[I], tmp2[I])
   res3 = res2
   res3[I] = tmp3[I]
return res2
\end{verbatim}
}
